Introduction to Moltres: an Application for Simulation of Molten Salt Reactors

### 2017-05-10 Moltres.tex
some rephrasings to put the spotlight on capability and scientific discovery

\gls{MSR} modeling efforts are focused on description of steady-state and
transient behavior. Krepel et. al. extended the in-house \gls{LWR}
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR \cite{krepel_dyn3d-msr_2007}. DYN3D-MSR was compared against
experimental \gls{MSRE} data and was then used to simulate local fuel channel
blockages as well as local temperature perturbations. In a similar vein, Kophazi
et. al. used iterative coupling between in-house three-dimensional neutronic and
one-dimensional heat conduction models DALTON and THERM to analyze the normal
operational state of the \gls{MSRE} as well as channel-blocking-incident
transients \cite{kophazi_development_2009}. Additions in the Kophazi model
included entrance effects of heat transfer coefficients as well as thermal
coupling between fuel channels through moderator heat conduction. More recently,
Cammi et. al. performed a 2D-axisymmetric single-channel analysis of the
\gls{MSBR} using the commercial finite element package COMSOL
Multiphysics \cite{cammi_multi-physics_2011}. Advances in this work included
solution of the fuel salt velocity field, use of heterogeneous group constants
in fuel and moderator regions, and employment of a software package (COMSOL)
intrinsically designed for coupled multi-physics simulation.

This work does not introduce new science relative to the current literature so
much as it introduces a new open tool, Moltres
\cite{lindsay_arfc/moltres_nodate}, for \gls{MSR} simulation. Moltres is an
open source code licensed under \gls{LGPL} terms that aims to be a tool that can be
freely used, studied, critiqued, and developed by the \gls{MSR}
community. Though Moltres is devoted to an accurate description of \gls{MSR}
physics, its openness is its defining characteristic and one that actually
promotes the former quality through transparency and easy peer review. In the
era of github and global development, open and modern software practices must be
employed in order for nuclear engineering simulation capability to keep pace
with that of fields like astronomy whose multitude of open simulation tools
(Enzo, Gadget, Flash, Gizmo, Ramses, Mesa) allows them to be the darlings of
scientific visualization and documentaries. In that vein Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another \gls{LGPL}
code that itself leans on LibMesh \cite{kirk2006libmesh}, a \gls{LGPL} finite element
library, and PetSc \cite{petsc-user-ref}, a \gls{BSD}-licensed toolkit for solving nonlinear
equations stemming from discretization of PDEs. \gls{MOOSE} and LibMesh handle
translation of application (Moltres) defined weak PDE forms into residual and
Jacobian functions that are the inputs into PetSc Newton-Raphson solution
routines. All codes use MPI for parallel communication and are easily deployed
on massively-parallel cluster-computing platforms. \gls{MOOSE} applications by default
use monolithic and implicit methods that are ideal for closely-coupled and
multi-scale physics. These methods are ideal for the model problem described in
this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it more
extensible to future problems.

>>>>

\gls{MSR} modeling efforts are focused on description of steady-state and
transient behavior. Krepel et. al. extended the in-house \gls{LWR}
diffusion code DYN3D to consider drift of delayed neutron precursors alongside
the reactor temperature profile, re-casting the extended code as
DYN3D-MSR \cite{krepel_dyn3d-msr_2007}. DYN3D-MSR was compared against
experimental \gls{MSRE} data and was then used to simulate local fuel channel
blockages as well as local temperature perturbations. In a similar vein, Kophazi
et. al. used iterative coupling between in-house three-dimensional neutronic and
one-dimensional heat conduction models DALTON and THERM to analyze the normal
operational state of the \gls{MSRE} as well as channel-blocking-incident
transients \cite{kophazi_development_2009}. Additions in the Kophazi model
included entrance effects of heat transfer coefficients as well as thermal
coupling between fuel channels through moderator heat conduction. More recently,
Cammi et. al. performed a 2D-axisymmetric single-channel analysis of the
\gls{MSBR} using the commercial finite element package COMSOL
Multiphysics \cite{cammi_multi-physics_2011}. Advances in that work included
solution of the fuel salt velocity field, use of heterogeneous group constants
in fuel and moderator regions, and employment of a software package (COMSOL)
intrinsically designed for coupled multi-physics simulation.

The present work introduces the open source simulation tool, Moltres 
\cite{lindsay_arfc/moltres_nodate}, for \gls{MSR} simulation.  By implementing 
deterministic neutronics and thermal hydraulics in the context of the 
\gls{MOOSE} finite element modeling framework, Moltres solves arbitrary-group 
neutron diffusion, temperature, and precursor governing equations in anywhere 
from one to three dimensions and can be deployed on an arbitrary number of 
processing units. Moltres is devoted to previously unmatched 
fidelity in coupled neutronics and thermal hydraulics \gls{MSR} simulation.

Moltres is an open source code licensed under \gls{LGPL} terms that aims to be a tool that can be
freely used, studied, interrogated, and improved by the \gls{MSR}
community.  Its openness is a defining characteristic and 
promotes unmatched quality through transparency and ease of peer review. In the
era of GitHub\cite{github} and international scientific collaboration, open and modern software practices must be
employed in order for nuclear engineering simulation capability to enable 
discovery and support the regulatory needs presented by new reactor designs. 
Furthermore, to openness is necessary to keep pace with scientific computing 
advances in fields like astronomy whose multitude of open simulation tools
(Enzo, Gadget, Flash, Gizmo, Ramses, Mesa) allows them to be the darlings of
scientific visualization and public interest. In that vein Moltres uses
\mintinline{latex}{git} for version control, integration testing to protect
developed physics capabilities, and a C++ object-oriented design to allow easy
class extensibility and minimal code duplication. Moreover, Moltres is built on
top of the \gls{MOOSE} framework, \cite{gaston_physics-based_2015} another \gls{LGPL}
code that itself leans on LibMesh \cite{kirk2006libmesh}, a \gls{LGPL} finite element
library, and PetSc \cite{petsc-user-ref}, a \gls{BSD}-licensed toolkit for solving nonlinear
equations stemming from discretization of PDEs. \gls{MOOSE} and LibMesh handle
translation of application (Moltres) defined weak PDE forms into residual and
Jacobian functions that are the inputs into PetSc Newton-Raphson solution
routines. All codes use MPI for parallel communication and are easily deployed
on massively-parallel cluster-computing platforms. \gls{MOOSE} applications by default
use monolithic and implicit methods that are ideal for closely-coupled and
multi-scale physics. These methods are ideal for the model problem described in
this work. However, it should be noted that Moltres can also use explicit
time-stepping routines as well as segregated solution methods, making it more
extensible to future problems.
